### 2020-04-19 paper/article.tex
more

% The MIT License (MIT)
%
% Copyright (c) 2020 Yegor Bugayenko
%
% Permission is hereby granted, free of charge, to any person obtaining a copy
% of this software and associated documentation files (the "Software"), to deal
% in the Software without restriction, including without limitation the rights
% to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
% copies of the Software, and to permit persons to whom the Software is
% furnished to do so, subject to the following conditions:
%
% The above copyright notice and this permission notice shall be included
% in all copies or substantial portions of the Software.
%
% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
% IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
% FITNESS FOR A PARTICULAR PURPOSE AND NON-INFRINGEMENT. IN NO EVENT SHALL THE
% AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
% LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
% OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
% SOFTWARE.

\documentclass[sigconf]{acmart}
\title{Volatility Metric to Detect Anomalies in Source Code Repositories}
\author{Yegor Bugayenko}{}{}
\email{yegor.bugayenko@huawei.com}
\affiliation{%
  \institution{Huawei Technologies Co., Ltd.}
  \city{Moscow, Russia}
}
\ccsdesc[100]{Software and its engineering~Metrics}
\keywords{Metrics, Software Size, Software Maintainability}
\setcopyright{acmcopyright}
\copyrightyear{2020}
\acmYear{2020}

\usepackage[utf8]{inputenc}
\usepackage{textcomp}
\usepackage[inline]{enumitem}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{verbatimbox}
\usepackage{interval}
\usepackage{hyperref}
\usepackage{minted}
  \setminted{fontsize=\footnotesize}
  \setminted{breaklines}
  \usemintedstyle{bw}
\newcommand{\code}[1]{\texttt{#1}}
\newenvironment{nicetable}
  {\setlength{\parindent}{0em}\medskip\small}
  {\medskip}
\input{total}
\begin{document}
\begin{abstract}
A new metric was introduced to calculate the distance
between actively modified files in a source code repository
and the files, which are rarely modified and may be considered
as ``dead'' or ``abandoned.'' It was empirically demonstrated that larger repositories
have larger values of the introduced metric.
The metric may be used for earlier detection of code maintenance anomalies
and helping software developers make the decision of splitting the repository
into smaller ones in order to prevent maintainability issues.
\end{abstract}
\maketitle

\section{Introduction}

Most software development projects keep their source code in Git~\citep{loeliger2012},
which is the de-facto standard in the industry at the moment, or similar
version control systems. Every system, including open
source products like Git, Subversion, and Mercurial, and commerical tools
like Borland StarTeam\texttrademark{} or IBM ClearCase\texttrademark{}
have the same feature: keeping track of the changes
made to the source code files, also known as ``logs.''

Git logs provide information about every single change made by every software
developer during the entire course of the project. Using this information
it's possible to measure which files are being modified frequently. On the
other hand, it's also possible to spot files that are rarely modified and may
be considered as ``dead'' or ``abandoned'' code, which may be considered as a threat
to the maintainability of the entire project.

We introduce a metric to measure the relationship
between the amount of actively modified files and files which stay
in the repository for a long time without modifications.

In order to demonstrate how the metric works we applied it to
\thetotalrepos{} public Java repositories from GitHub and analyzed
its impact on a two other characteristics of each repository:
the number of files in a repository and the size of it in bytes.
Larger repositories shown larger values of the volatility.

The paper is organized as follows.
Section~\ref{sec:background} defines various terms used in the paper.
Section~\ref{sec:related} covers related work in the areas of code change metrics
and empirical analysis of source code repository size.
Section~\ref{sec:method} covers the source code volatility metric.
Section~\ref{sec:results} covers our empirical case study of \thetotalrepos{} Java source code repositories.
Section~\ref{sec:discussion} covers limitations of both the metric and the study.
Finally, we summarize our study in Section~\ref{sec:conclusion}.

\section{Background}
\label{sec:background}

In this paper, we use several terms regarding version control systems and
software metrics. The type of metric we are proposing belongs to the family
of source code change metrics, which analyze the history
of changes in a source code repository~\citep{fentonsoftware,choudhary2018}.
A \emph{change} in a version control system, such as Git, is an individual
modification to the source code file made by a software developer locally
on his/her machine and then \emph{committed} to the repository.

Each version control system, including Git, provides an instrument for
retrieving the entire history of changes from a repository. We use
\texttt{git log} for Git.

GitHub\texttrademark{} (currently owned by Microsoft\texttrademark{}) is
one of the largest platforms for \emph{open source} projects.
GitHub provides free access to all public Git repositories, which makes it possible
to analyze Git histories.

\section{Related Work}
\label{sec:related}

``Code change metrics mined from source control repositories have
proven to be the most reliable predictors of bugs in
contemporary software engineering research,'' says \citet{muthukumaran2015}.

Code churn is one of the simplest metric in the family, which counts
all lines being modified (added, edited, and deleted) during
the entire lifetime of a project~\citep{munson1998}. It was demonstrated
by~\citet{shin2010} how code churn, together with complexity and other
developer activity metrics, is related to software vulnerabilities.
Yet another simple metrict from the family is
the number of changes being made to a particular file, class, method
or line of code is yet another simple metric, used for example by~\citet{demeyer2000}
in order to analyze the effect of refactorings.

\citet{moser2008} made a comparative analysis of 17 change metrics to understand
their efficiency for defect prediction: number of distinct authors per file,
total modifications per file, total additions per file, maximum number of
files committed together, and others.

\citet{biazzini2014} introduced a number of metrics to analyze
Git history (specifically related to GitHub), such as unique-count, unique-ratio,
VIP-count, VIP-ratio, scattered-count, pervasive-count and a few others. Some
of the metrics are supposed to be calculated for a repository together with
its forks, while others may be calculated for a single repository history.

\citet{batista2018} provided a detailed analysis of existing metrics
and introduced a new one to measure, by looking at Git/GitHub project
commit history, how ``close'' developers stay to each other and form
pairs.

To our knowledge, the metric introduced in this research has never been
suggested.

\section{Metric}
\label{sec:method}

First, by looking at the Git history,
it is observed how many times every source code file out of $N$ was touched
during the lifetime of the repository (excluding the files that don't exist
in the repository anymore):

\begin{eqnarray}
T = [t_1, t_2, \dots, t_N]
\end{eqnarray}

Then, the entire interval between $\check{T}$ (the maximum value)
and $\hat{T}$ (the minimum value) is divided to $Z$ (the input parameter of the method)
equivalent groups:

\begin{align}
\delta &= ( \check{T} - \hat{T} ) / Z \\
G &= [g_1, g_2, \dots, g_{Z}] \\
g_j &= \sum_{i=1}^N [ j(\delta-1) < t_i < j\delta ]
\end{align}

Then, the mean $\mu$ is calculated as:

\begin{eqnarray}
\mu = \frac{1}{Z}\sum_{j=1}^{Z}{g_j}
\end{eqnarray}

Finally, the variance is calculated as:

\begin{eqnarray}
\text{Var}(g) = \frac{1}{Z}\sum_{j=1}^{Z}{|g_j - \mu|^2}
\end{eqnarray}

The variance $\text{Var}(g)$ is the volatility of the source code. The smaller
the volatility the more cohesive is the repository and the smaller
the amount of the dead code inside it.

\section{Empirical Results}
\label{sec:results}

A list of Java repositories were retrieved from GitHub via their
public API. The first \thetotalrepos{} repositories were taken, which satisfied
the selection criteria:
\begin{enumerate*}[label={\arabic*)}]
\item more than 1,000 GitHub stars,
\item more than 200 Kb of data,
\item not archived, and
\item public.
\end{enumerate*}
The list included popular Java open source products, such as
Spring, RxJava, Guava, MyBatis, Clojure, JUnit, Lombok,
Graal, Selenium, Spark, Mockito, Neo4j, Jenkins, Netty, and others.

The volatility metric was calculated for each repository, using the
formula explained above (the value of $Z$ was set to 64).
Then, a few other metrics were collected
for each repository and their values were compared with the volatility.

The Figure~\ref{fig:1} demonstrates the relationship between
the number of files in the repository~($M_1$) and its volatility~($V$). Both
axixes of the graph have logarithmic scales, for the sake of visual
understandability: the difference between minimum and maximum values
of volatility is logarithmically large. It is visually obvious that
repositories with larger number of files tend to have higher values
of the volatility metric.

\begin{figure}[h]
  \input{files.tex}
  \caption{The relationship between the number of files in the repository and its volatility}
  \label{fig:1}
\end{figure}

The Figure~\ref{fig:2} demonstrates the relationship between
the logarithm of the size of the Git repository in bytes~($M_2$) and
the logarithm of its volatility~($V$).
It is visually obvious that
binary-size larger repositories tend to have higher values
of the volatility metric.

\begin{figure}[h]
  \input{bytes.tex}
  \caption{The relationship between the size of the Git repository in bytes and its volatility}
  \label{fig:2}
\end{figure}

\section{Discussion}
\label{sec:discussion}

It was demonstrated that the size of the source code is an important
factor negatively affecting maintainability and other quality characteristicts.
For example, \citet{heitlager2007} suggested to consider ``volume'' (the amount of code
in the repository, the number of lines per file, the number of methods in a class,
and so on) as a primary contributor to an aggregate maintainability index
of a software because ``it is fairly intuitive that the total size of a system should
feature heavily in any measure of maintainability.''

\section{Conclusion}
\label{sec:conclusion}

A new source code volatility metric was introduced and applied
to \thetotalrepos{} open source Java repositories from GitHub. It was
empirically demonstrated that larger repositories have higher values
of the volatility metric.

% The source code of Ruby and Python scripts used to do the research
% is available in GitHub repository \texttt{yegor256/volatility-vs-size}.

\bibliographystyle{ACM-Reference-Format}
\bibliography{main}

\end{document}